\documentclass[12pt]{report}

\usepackage[a4paper, margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{graphicx}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    citecolor=black,
    bookmarksdepth=section
}

\setlength{\headheight}{13.6pt}
\addtolength{\topmargin}{-1.6pt}

\lstset{
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    backgroundcolor=\color{gray!5},
    rulecolor=\color{gray!40},
    xleftmargin=2em,
    framexleftmargin=1.5em,
    numbers=none,
    tabsize=4,
    showstringspaces=false
}

\definecolor{warnorange}{RGB}{204,102,0}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small Blentinel Technical Manual}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\title{Blentinel Monitoring Platform\\
\large Technical Manual}
\author{
    \textbf{James Seaton Reid}\\
    \textit{Lead Software Architect}
}
\date{Version 0.1.0 \quad---\quad \today}

\begin{document}

\maketitle
\tableofcontents
\newpage

% ============================================================
% FRONT MATTER
% ============================================================

\chapter*{Project Overview}
\addcontentsline{toc}{chapter}{Project Overview}

Blentinel, the \textbf{Bl}ind S\textbf{entinel}, is a high-performance,
multi-tenant monitoring platform designed for Managed Service Providers
(MSPs). It enables a central IT organisation to monitor distributed
client networks, including those behind NAT or private firewalls,
using a secure push-based architecture.

The system is composed of a \textbf{Hub} (the server) and one or more \textbf{Probes} (clients).
Probes operate inside client environments and push signed, encrypted telemetry
to the Hub at configurable intervals. The Hub authenticates every report,
decrypts it, stores the data in SQLite, and presents monitoring information
through a server-rendered web interface.

The platform is written entirely in Rust and compiles to compact, statically
linkable binaries for both Linux and Windows.

\chapter*{Authors}
\addcontentsline{toc}{chapter}{Authors}

\begin{longtable}{ll}
\toprule
\textbf{Name} & \textbf{Role} \\
\midrule
James Seaton Reid & Lead Software Architect \\
Open AI ChatGPT & Technical Advisor \\
Anthropic Claude Code & Lead Developer \\

\bottomrule
\end{longtable}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART I — OPERATOR MANUAL
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{Operator Manual}
\setcounter{section}{0}

% ============================================================
\section{Motivation and Design Philosophy}
% ============================================================

Traditional monitoring solutions typically operate on a private network and
in many cases need for the networks they monitor be visible to the server,
require inbound firewall rules, VPN connectivity, or persistent network tunnels.
These approaches increase attack surface and operational complexity, especially
for MSPs managing dozens of client networks.

Blentinel is built around a strict \textbf{outbound-only} model:

\begin{itemize}
    \item Probes initiate all connections.
    \item No inbound firewall rules are required on client networks.
    \item The Hub never connects back to probes.
\end{itemize}

This model improves security posture and simplifies deployment in zero-trust
and MSP environments.

\subsection{Key Design Principles}

\begin{enumerate}
    \item \textbf{Probe zero dependency policy} --- As probes are designed to
    	  be deployed on remote private networks they must have zero dependencies
    	  and no runtimes.
    \item \textbf{Push-based communication} --- probes push data; the hub is
		  purely passive.
    \item \textbf{Cryptographic identity per probe} --- every probe holds a
          unique Ed25519 keypair.
    \item \textbf{Explicit probe whitelisting} --- the hub rejects any probe
          not registered in its configuration.
    \item \textbf{Defence-in-depth encryption} --- application-layer
          authenticated encryption (X25519 + ChaCha20-Poly1305) with optional
          TLS on the transport layer.
    \item \textbf{Perfect forward secrecy} --- ephemeral X25519 keys per
          report ensure past reports remain confidential even if long-term keys
          are compromised.
    \item \textbf{Minimal operational complexity} --- single-binary deployment,
          TOML configuration, automatic key and certificate generation.
    \item \textbf{Store-and-forward resilience} --- probes queue encrypted
          reports locally when the hub is unreachable (BlackBox).
\end{enumerate}

% ============================================================
\section{System Architecture}
% ============================================================

Blentinel consists of four workspace crates:

\subsection{Common Library (\texttt{common})}

Shared data models used by both probe and hub, including:

\begin{itemize}
    \item \texttt{StatusReport} --- the canonical report structure.
    \item \texttt{SignableReport} --- the subset of fields that are
          cryptographically signed (excludes signature and ephemeral key).
    \item \texttt{ResourceStatus} --- a single resource's health snapshot.
    \item \texttt{ResourceType} and \texttt{Health} enumerations.
\end{itemize}

\subsection{Sentinel Probe (\texttt{probe})}

A low-footprint service deployed inside monitored networks. It:

\begin{itemize}
    \item Polls local and network resources at a configurable interval.
    \item Signs each report with Ed25519.
    \item Encrypts the report with X25519 + ChaCha20-Poly1305.
    \item Pushes the encrypted payload to the hub via HTTP(S).
    \item Queues reports in a local BlackBox database when the hub is
          unreachable, replaying them automatically once connectivity is
          restored.
    \item Supports hot-reloading of its configuration file.
\end{itemize}

\subsection{Blind Sentinel Hub (\texttt{hub})}

The central coordination service responsible for:

\begin{itemize}
    \item Authenticating probe identities via whitelist lookup.
    \item Decrypting incoming payloads using its persistent X25519 private key.
    \item Verifying Ed25519 digital signatures.
    \item Storing time-series data in SQLite (WAL mode).
    \item Tracking probe liveness (active/expired).
    \item Serving a full-stack Leptos web interface (server-side rendered with
          WebAssembly hydration).
    \item Managing data retention and archival.
    \item Sending e-mail alerts on resource failures, threshold breaches, and
          probe expiry.
    \item Hot-reloading probe whitelist and timeout settings.
\end{itemize}

\subsection{Build Tool (\texttt{blentinelmake})}

A cross-platform build and packaging utility that replaces ad-hoc scripts.
Provides both an interactive terminal menu and a CLI interface for building,
publishing, and cleaning hub and probe artefacts. Generates SHA-256 integrity
checksums for published binaries.

% ============================================================
\section{Technology Stack}
% ============================================================

\begin{longtable}{@{} l l @{}}
\toprule
\textbf{Component} & \textbf{Technologies} \\
\midrule
Probe         & Rust (Tokio, reqwest, surge-ping, sysinfo) \\
Hub           & Rust (Tokio, Axum, Leptos 0.8, SQLx) \\
Database      & SQLite (WAL mode) \\
Frontend      & Leptos (SSR + WASM hydration), SCSS \\
Transport     & Optional TLS/HTTPS (Rustls, axum-server) \\
Payload Encryption & X25519 key exchange + ChaCha20-Poly1305 AEAD \\
Authentication     & Ed25519 digital signatures \\
Certificate Gen.   & rcgen (ECDSA P-256, self-signed) \\
Alerting      & lettre (SMTP) \\
Build Tool    & Rust (dialoguer, sha2) \\
\bottomrule
\end{longtable}

% ============================================================
\section{Secure Bootstrap Procedure}
% ============================================================

This section describes the correct first-time setup sequence.

\subsection{Step 1 --- Build and Deploy Hub}

\begin{lstlisting}
cargo build -p blentinelmake --release
./target/release/blentinelmake hub publish
\end{lstlisting}

Deploy the generated \texttt{publish/hub/app/} directory to the server.

Initialise configuration:

\begin{lstlisting}
./hub --init
\end{lstlisting}

Review and edit \texttt{blentinel\_hub.toml}, then start the hub:

\begin{lstlisting}
./hub
\end{lstlisting}

On first run, the hub generates:

\begin{itemize}
    \item \texttt{hub\_identity.key} --- persistent X25519 private key (32
          bytes).
    \item \texttt{hub\_auth.token} --- admin authentication token (32-char
          hex). This token is printed to the terminal on first run.
    \item \texttt{blentinel.db} --- SQLite database (created automatically).
\end{itemize}

\textcolor{warnorange}{\textbf{Warning:}} The identity key and auth token
files must be protected. Loss of the identity key invalidates all queued
BlackBox reports across all probes.

\subsection{Step 2 --- Build and Deploy Probe}

\begin{lstlisting}
./target/release/blentinelmake probe publish \
    --target x86_64-unknown-linux-gnu
\end{lstlisting}

Deploy the \texttt{publish/probe/<target>/app/} directory to the probe host.

Initialise configuration:

\begin{lstlisting}
./probe --init
\end{lstlisting}

Edit \texttt{blentinel\_probe.toml} to set the company ID, hub URL, monitoring
interval, and resources.

\subsection{Step 3 --- Probe Identity Generation}

On first run, the probe generates an Ed25519 keypair and prints its public key:

\begin{lstlisting}
./probe
\end{lstlisting}

The public key is saved to \texttt{identity.key} (32-byte private key). The
public key (hex-encoded) is displayed on the terminal and serves as the
probe's unique identifier.

\subsection{Step 4 --- Register Probe in Hub}

Add the probe's public key to \texttt{blentinel\_hub.toml}:

\begin{lstlisting}
[[probes]]
name = "OFFICE-PC-1"
public_key = "23729607766e0ddce6d88f...fbc4d5"
\end{lstlisting}

The hub hot-reloads this change automatically --- no restart required.

\subsection{Step 5 --- Handshake}

On startup, if the probe does not have the hub's public key cached, it
performs a handshake:

\begin{enumerate}
    \item Probe sends \texttt{GET /api/handshake} to the hub.
    \item Hub returns its X25519 public key (hex-encoded).
    \item Probe caches the key in memory for subsequent reports.
    \item If \texttt{hub\_public\_key} is set in the probe's configuration,
          the probe verifies the received key matches.
\end{enumerate}

The probe retries the handshake every 120 seconds if the hub is unreachable.

\subsection{Correct Bootstrap Order}

\begin{enumerate}
    \item Start the Hub.
    \item Start the Probe (generates its keypair on first run).
    \item Copy the probe's public key and add it to the hub configuration.
    \item The hub hot-reloads the whitelist; the probe's next report is
          accepted.
\end{enumerate}

% ============================================================
\section{Service Installation}
% ============================================================

\subsection{Linux (systemd)}

Create the service file at
\texttt{/etc/systemd/system/blentinel-probe.service}:

\begin{lstlisting}
[Unit]
Description=Blentinel Probe
After=network.target

[Service]
ExecStart=/opt/blentinel/probe/probe
WorkingDirectory=/opt/blentinel/probe
Restart=always
User=blentinel

[Install]
WantedBy=multi-user.target
\end{lstlisting}

Enable and start:

\begin{lstlisting}
sudo systemctl enable blentinel-probe
sudo systemctl start blentinel-probe
\end{lstlisting}

For the hub, create an analogous \texttt{blentinel-hub.service} with the
appropriate paths.

\textbf{Note on ICMP:} The probe requires raw socket permissions for ping
checks. Either run as root or grant the \texttt{CAP\_NET\_RAW} capability:

\begin{lstlisting}
sudo setcap cap_net_raw+ep /opt/blentinel/probe/probe
\end{lstlisting}

\subsection{Windows}

Use the PowerShell installer script included in the published package, or
install manually using \texttt{sc.exe} or a third-party service wrapper such
as NSSM.

% ============================================================
\section{Configuration Reference}
% ============================================================

\subsection{Probe Configuration (\texttt{blentinel\_probe.toml})}

\subsubsection{Agent Section}

\begin{longtable}{@{} l l p{7cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{company\_id} & String & Unique company identifier. Must match across
    all probes belonging to the same client. \\
\texttt{hub\_url} & String & Full URL of the hub (e.g.\
    \texttt{http://10.0.0.1:3000}). Use \texttt{https://} when TLS is
    enabled. \\
\texttt{interval} & Integer & Reporting interval in seconds. Must be greater
    than 0. \\
\texttt{hub\_public\_key} & String? & Optional hex-encoded X25519 public key
    of the hub. When set, the probe verifies the handshake response matches
    this value. \\
\texttt{site} & String & Logical site name (default: \texttt{"main"}). Used
    to group probes within a company. \\
\bottomrule
\end{longtable}

\subsubsection{Resources Section}

Each \texttt{[[resources]]} entry defines a single monitored target:

\begin{longtable}{@{} l l p{7cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{name} & String & Human-readable name for this resource. \\
\texttt{type} & String & Resource type (see Section~\ref{sec:resource-types}).
    \\
\texttt{target} & String & Target address. Format depends on the type. \\
\bottomrule
\end{longtable}

\subsubsection{Resource Types}
\label{sec:resource-types}

\begin{longtable}{@{} l p{3cm} p{7cm} @{}}
\toprule
\textbf{Type} & \textbf{Target Format} & \textbf{Description} \\
\midrule
\texttt{ping} & IP address & ICMP echo request. Reports latency in
    milliseconds. Requires raw socket permissions on Linux. \\
\texttt{http} & URL & HTTP(S) GET request. Reports status code. A 2xx
    response is considered healthy; any other code or connection failure is
    ``Down''. \\
\texttt{tcp} & \texttt{host:port} & TCP connection attempt with 3-second
    timeout. Reports whether the port is open and connection latency. \\
\texttt{local\_disk} & Mount point or \texttt{auto} & Disk usage for the
    specified mount point (e.g.\ \texttt{/} or \texttt{C:\textbackslash}).
    Setting \texttt{auto} or an empty string selects the system drive.
    Reports percentage used. \\
\texttt{local\_cpu} & \texttt{localhost} & Global CPU usage percentage. Uses
    \texttt{sysinfo} for cross-platform support. \\
\texttt{local\_mem} & \texttt{localhost} & Memory usage. Reports used/total in
    GB and percentage. \\
\texttt{local\_uptime} & \texttt{localhost} & System uptime in days, hours,
    and minutes. The \texttt{metric\_value} field contains uptime in seconds.
    \\
\texttt{local\_data} & \texttt{localhost} & \textbf{Meta-type.} Expands into
    four checks: CPU, memory, disk (system drive), and uptime. Resource names
    are suffixed, e.g.\ ``Server (CPU)'', ``Server (Memory)''. \\
\bottomrule
\end{longtable}

\subsubsection{Example Probe Configuration}

\begin{lstlisting}
[agent]
company_id = "IT_CLIENT_ALPHA"
hub_url = "http://10.0.0.1:3000"
interval = 60
site = "main"
# hub_public_key = "c746...b071"

[[resources]]
name = "Router"
type = "ping"
target = "192.168.1.1"

[[resources]]
name = "NAS Web UI"
type = "http"
target = "https://192.168.1.224"

[[resources]]
name = "Database"
type = "tcp"
target = "192.168.1.50:5432"

[[resources]]
name = "System Disk"
type = "local_disk"
target = "/"

[[resources]]
name = "Server Health"
type = "local_data"
target = "localhost"
\end{lstlisting}

\subsection{Hub Configuration (\texttt{blentinel\_hub.toml})}

\subsubsection{Server Section}

\begin{longtable}{@{} l l l p{5.5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{host} & String & --- & Bind address (e.g.\ \texttt{0.0.0.0}). \\
\texttt{port} & Integer & --- & HTTP listen port. \\
\texttt{db\_path} & String & --- & Path to the SQLite database file. \\
\texttt{identity\_key\_path} & String & \texttt{hub\_identity.key} &
    Path to the persistent X25519 private key file. \\
\texttt{auth\_token\_path} & String & \texttt{hub\_auth.token} &
    Path to the admin authentication token file. \\
\texttt{probe\_timeout\_secs} & Integer & \texttt{300} &
    Seconds of silence before a probe is marked expired. Should be 2--3$\times$
    the probe's reporting interval. Hot-reloadable. \\
\bottomrule
\end{longtable}

\subsubsection{TLS Section (\texttt{[server.tls]})}

\begin{longtable}{@{} l l l p{5.5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{enabled} & Boolean & \texttt{false} & Enable TLS/HTTPS. \\
\texttt{cert\_path} & String & \texttt{hub\_tls\_cert.pem} &
    Path to the TLS certificate (PEM). \\
\texttt{key\_path} & String & \texttt{hub\_tls\_key.pem} &
    Path to the TLS private key (PEM). \\
\texttt{https\_port} & Integer? & \texttt{None} &
    If set, the hub listens on both HTTP (\texttt{port}) and HTTPS
    (\texttt{https\_port}). If omitted, HTTPS replaces HTTP on the main port.
    \\
\bottomrule
\end{longtable}

\subsubsection{Probes Section (\texttt{[[probes]]})}

Each entry registers an authorised probe:

\begin{longtable}{@{} l l p{7cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{name} & String & Human-readable probe name. \\
\texttt{public\_key} & String & Hex-encoded Ed25519 public key (64
    hex characters = 32 bytes). \\
\bottomrule
\end{longtable}

\subsubsection{Retention Section (\texttt{[retention]})}

\begin{longtable}{@{} l l l p{4.5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{enabled} & Boolean & \texttt{true} & Enable retention monitoring.
    \\
\texttt{auto} & Boolean & \texttt{false} & Automatically archive old
    data when the database exceeds \texttt{warn\_db\_size\_mb}. \\
\texttt{archive\_older\_than\_days} & Integer & \texttt{90} & Data older
    than this many days is eligible for archival. \\
\texttt{warn\_db\_size\_mb} & Integer & \texttt{1000} & Database size
    warning threshold in megabytes. \\
\texttt{archive\_path} & String & \texttt{archives} & Directory where
    archive databases are stored. \\
\bottomrule
\end{longtable}

\subsubsection{Alerts Section (\texttt{[alerts]})}

\begin{longtable}{@{} l l l p{4.5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{enabled} & Boolean & \texttt{false} & Enable the alerting system. \\
\texttt{default\_recipients} & [String] & \texttt{[]} & Default e-mail
    recipients for all alerts. \\
\bottomrule
\end{longtable}

\paragraph{Technicians (\texttt{[[alerts.technicians]]})}

\begin{longtable}{@{} l l p{7cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{name} & String & Technician's full name. \\
\texttt{email} & String & E-mail address. \\
\texttt{phone} & String? & Optional phone number. \\
\bottomrule
\end{longtable}

\paragraph{Thresholds (\texttt{[alerts.thresholds]})}

\begin{longtable}{@{} l l l p{5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{disk\_percent} & Integer & \texttt{90} & Alert when disk usage exceeds
    this percentage. \\
\texttt{cpu\_percent} & Integer & \texttt{95} & Alert when CPU usage exceeds
    this percentage. \\
\texttt{mem\_percent} & Integer & \texttt{90} & Alert when memory usage exceeds
    this percentage. \\
\bottomrule
\end{longtable}

\paragraph{SMTP (\texttt{[alerts.smtp]})}

\begin{longtable}{@{} l l l p{5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Default} & \textbf{Description} \\
\midrule
\texttt{server} & String & \texttt{""} & SMTP server hostname. Required when
    alerts are enabled. \\
\texttt{port} & Integer & \texttt{587} & SMTP port (587 for STARTTLS). \\
\texttt{username} & String & \texttt{""} & SMTP authentication username. \\
\texttt{password} & String & \texttt{""} & SMTP authentication password. \\
\texttt{from} & String & \texttt{""} & Sender address for alert e-mails.
    Required when alerts are enabled. \\
\bottomrule
\end{longtable}

\paragraph{Per-Company Overrides (\texttt{[alerts.company\_overrides."ID"]})}

Override alert recipients and thresholds for a specific company:

\begin{lstlisting}
[alerts.company_overrides."CompanyA"]
alert_emails = ["ops@companya.com"]

[alerts.company_overrides."CompanyA".thresholds]
disk_percent = 85
cpu_percent = 90
mem_percent = 85
\end{lstlisting}

\subsubsection{Example Hub Configuration}

\begin{lstlisting}
[server]
host = "0.0.0.0"
port = 3000
db_path = "blentinel.db"
identity_key_path = "hub_identity.key"
auth_token_path = "hub_auth.token"
probe_timeout_secs = 120

# [server.tls]
# enabled = true
# cert_path = "hub_tls_cert.pem"
# key_path = "hub_tls_key.pem"
# https_port = 3443

[[probes]]
name = "Office-PC-1"
public_key = "23729607...fbc4d5"

[retention]
enabled = true
auto = false
archive_older_than_days = 90
warn_db_size_mb = 1000
archive_path = "archives"

[alerts]
enabled = false
default_recipients = []

[alerts.thresholds]
disk_percent = 90
cpu_percent = 95
mem_percent = 90

[alerts.smtp]
server = "smtp.example.com"
port = 587
username = "alerts@example.com"
password = "your-smtp-password"
from = "blentinel@example.com"
\end{lstlisting}

% ============================================================
\section{HTTPS Configuration}
% ============================================================

Blentinel supports optional TLS/HTTPS with certificate pinning.

\subsection{Certificate Generation}

When TLS is enabled and certificate files do not exist, the hub
automatically generates:

\begin{itemize}
    \item A self-signed ECDSA P-256 certificate valid for 10 years.
    \item Subject Alternative Names: the configured hostname,
          \texttt{localhost}, \texttt{127.0.0.1}, \texttt{::1}.
    \item The SHA-256 certificate fingerprint is printed to the terminal.
\end{itemize}

\subsection{Operating Modes}

\paragraph{HTTP Only (Default)}

\begin{lstlisting}
[server.tls]
enabled = false
\end{lstlisting}

The hub listens on a single HTTP port.

\paragraph{HTTPS Only}

\begin{lstlisting}
[server.tls]
enabled = true
\end{lstlisting}

HTTPS replaces HTTP on the main port (no \texttt{https\_port} specified).

\paragraph{Dual Mode (Recommended for Migration)}

\begin{lstlisting}
[server.tls]
enabled = true
https_port = 3443
\end{lstlisting}

The hub listens on both HTTP (\texttt{port}) and HTTPS
(\texttt{https\_port}), allowing gradual probe migration.

\subsection{Certificate Pinning}

Probe-side certificate pinning works as follows:

\begin{enumerate}
    \item Copy \texttt{hub\_tls\_cert.pem} from the hub into the probe's build
          directory as \texttt{hub\_cert.pem}.
    \item Rebuild the probe: the certificate is embedded at compile time via
          \texttt{include\_bytes!}.
    \item Update the probe's \texttt{hub\_url} to use \texttt{https://}.
\end{enumerate}

\textbf{Important:} Certificate changes require a probe rebuild. This is by
design --- it ensures probes only trust a specific certificate, preventing
man-in-the-middle attacks even if an attacker possesses valid CA certificates.

\subsection{Migration Path (HTTP to HTTPS)}

\begin{enumerate}
    \item Enable dual mode on the hub (\texttt{https\_port = 3443}).
    \item Restart the hub; note the certificate fingerprint.
    \item Copy \texttt{hub\_tls\_cert.pem} to each probe's build directory.
    \item Rebuild and redeploy probes with \texttt{hub\_url =
          "https://HUB:3443"}.
    \item Once all probes are migrated, remove \texttt{https\_port} to serve
          HTTPS only on the main port.
\end{enumerate}

% ============================================================
\section{Alerting System}
% ============================================================

The hub can send e-mail alerts for resource failures, threshold breaches, and
probe expiry events.

\subsection{Alert Types}

\begin{itemize}
    \item \textbf{Resource state change:} Sent when a resource transitions
          from Up to Down (failure) or Down to Up (recovery).
    \item \textbf{Threshold breach:} Sent when a local metric
          (CPU, memory, disk usage) exceeds the configured threshold.
    \item \textbf{Probe expiry:} Sent when a probe stops reporting and its
          status changes to ``expired''.
\end{itemize}

\subsection{Alert Silences}

Alerts can be temporarily or permanently suppressed at three scopes:

\begin{itemize}
    \item \textbf{Resource:} Silence a specific resource.
    \item \textbf{Probe:} Silence all resources for a probe.
    \item \textbf{Company:} Silence all resources for a company.
\end{itemize}

Silences can have an expiry time (in hours) or be indefinite. They are
managed via the API or the web admin panel.

\subsection{Per-Company Overrides}

Each company can have its own alert recipients and metric thresholds. When a
company override is configured, alerts for that company's resources are sent
to the override recipients instead of (or in addition to) the default
recipients.

% ============================================================
\section{Data Retention and Archival}
% ============================================================

The hub includes a built-in archival system to manage long-term data growth.

\subsection{How Archival Works}

\begin{enumerate}
    \item A cutoff date is calculated: current time minus
          \texttt{archive\_older\_than\_days}.
    \item All reports and resource statuses older than the cutoff are copied to
          a new SQLite archive database in the \texttt{archive\_path}
          directory.
    \item The archived data is deleted from the main database.
    \item An archive metadata record is inserted into the \texttt{archives}
          table.
\end{enumerate}

\subsection{Triggering Archival}

\begin{itemize}
    \item \textbf{Manual:} Via the admin panel or
          \texttt{POST /api/admin/archive}.
    \item \textbf{Automatic:} When \texttt{retention.auto = true}, the hub
          monitors database size every 10 minutes and triggers archival if the
          size exceeds \texttt{warn\_db\_size\_mb}.
\end{itemize}

\subsection{Browsing Archives}

Archived data can be browsed through the archive viewer API endpoints. The web
interface provides the same dashboard, company, probe, and uptime views for
historical archive databases.

% ============================================================
\section{Configuration Hot Reload}
% ============================================================

Both the hub and probe watch their configuration files for changes. When a
change is detected:

\begin{enumerate}
    \item The new file is read and parsed.
    \item Validation rules are applied (same rules as initial load).
    \item If valid, the new configuration replaces the old one atomically.
    \item If invalid, the error is logged and the previous configuration is
          retained.
\end{enumerate}

\subsection{Probe --- Fully Hot-Reloadable}

All probe settings are hot-reloadable:

\begin{itemize}
    \item Agent settings (interval, hub URL, company ID, site).
    \item Resource list (add, remove, or modify monitored resources).
\end{itemize}

\subsection{Hub --- Partially Hot-Reloadable}

\paragraph{Hot-reloadable (no restart required):}

\begin{itemize}
    \item Probe whitelist (\texttt{[[probes]]}).
    \item \texttt{probe\_timeout\_secs}.
\end{itemize}

\paragraph{Restart required:}

\begin{itemize}
    \item \texttt{server.host} and \texttt{server.port}.
    \item \texttt{db\_path}.
    \item \texttt{identity\_key\_path}.
    \item \texttt{auth\_token\_path}.
    \item TLS configuration (\texttt{[server.tls]}).
\end{itemize}

% ============================================================
\section{CLI Reference}
% ============================================================

\subsection{Probe}

\begin{lstlisting}
Usage: blentinel_probe [OPTIONS]

Options:
  -h, --help      Print help message and exit
  --version       Print compiled version and exit
  --init          Create default blentinel_probe.toml and exit
  -v, --verbose   Log operational events (handshake, delivery,
                  connectivity)
  -d, --debug     Log cleartext payloads before encryption
                  (implies --verbose; never use in production)
\end{lstlisting}

\subsection{Hub}

\begin{lstlisting}
Usage: blentinel_hub [OPTIONS]

Options:
  -h, --help      Print help message and exit
  --version       Print compiled version and exit
  --init          Create default blentinel_hub.toml and exit
  -v, --verbose   Log operational events (reports received,
                  data saved, probe expiry)
  -d, --debug     Verbose mode + detailed diagnostic output
                  (implies --verbose; never use in production)
\end{lstlisting}

\subsection{blentinelmake}

\begin{lstlisting}
Usage: blentinelmake [COMPONENT] [ACTION] [OPTIONS]

Components:
  hub              Leptos-based monitoring hub
  probe            Network monitoring probe

Actions:
  build            Build the component
  publish          Build release + package for distribution
  clean            Remove build artefacts

Options:
  --target TRIPLE  Target triple for cross-compilation
                   (probe only)
  --release        Build in release mode (build action only)
  -h, --help       Print help and exit
  --version        Print version and exit
\end{lstlisting}

Run without arguments for interactive mode.

% ============================================================
\section{Operational Behaviour}
% ============================================================

\subsection{Generated Runtime Files}

\paragraph{Hub:}

\begin{longtable}{@{} l p{8cm} @{}}
\toprule
\textbf{File} & \textbf{Description} \\
\midrule
\texttt{hub\_identity.key} & Persistent X25519 private key (32 bytes). \\
\texttt{hub\_auth.token} & Admin authentication token (32-char hex). \\
\texttt{hub\_tls\_cert.pem} & TLS certificate (if TLS enabled). \\
\texttt{hub\_tls\_key.pem} & TLS private key (if TLS enabled). \\
\texttt{blentinel.db} & SQLite database. \\
\texttt{archives/} & Directory containing archive databases. \\
\bottomrule
\end{longtable}

\paragraph{Probe:}

\begin{longtable}{@{} l p{8cm} @{}}
\toprule
\textbf{File} & \textbf{Description} \\
\midrule
\texttt{identity.key} & Ed25519 private key (32 bytes). \\
\texttt{blentinel\_blackbox.db} & Store-and-forward queue (SQLite). \\
\bottomrule
\end{longtable}

\subsection{BlackBox Store-and-Forward}

When the hub is unreachable, the probe queues encrypted reports in the local
BlackBox database. On subsequent cycles:

\begin{enumerate}
    \item The probe attempts to send the current report.
    \item If successful, it replays all queued BlackBox reports in order
          (oldest first).
    \item Each successfully delivered report is deleted from the BlackBox.
\end{enumerate}

This ensures no telemetry is lost during temporary network outages or hub
maintenance windows.

% ============================================================
\section{Troubleshooting}
% ============================================================

\subsection{Invalid Signature}

\textbf{Symptom:} Hub logs ``Security Alert: Invalid signature'' and returns
403.

\textbf{Cause:} The probe's Ed25519 public key in the hub whitelist does not
match the key used to sign the report.

\textbf{Resolution:}
\begin{enumerate}
    \item Verify the public key in \texttt{blentinel\_hub.toml} matches the
          probe's output from its first run.
    \item If the probe's \texttt{identity.key} was regenerated, update the hub
          whitelist with the new public key.
\end{enumerate}

\subsection{Decryption Failed}

\textbf{Symptom:} Hub logs ``Decryption failed'' and returns 401.

\textbf{Causes:}
\begin{itemize}
    \item Hub's \texttt{hub\_identity.key} was regenerated (new X25519 key).
    \item Probe has a stale hub public key cached or in BlackBox.
    \item Corrupted payload (network issue).
\end{itemize}

\textbf{Resolution:}
\begin{enumerate}
    \item If the hub key changed, all probes must perform a new handshake.
    \item Delete \texttt{blentinel\_blackbox.db} on the probe to force a clean
          state.
    \item If using \texttt{hub\_public\_key} in probe config, update it.
\end{enumerate}

\subsection{Probe Marked as Expired}

\textbf{Symptom:} Probe shows as ``expired'' in the dashboard despite being
running.

\textbf{Causes:}
\begin{itemize}
    \item \texttt{probe\_timeout\_secs} is too low relative to the probe's
          interval. Set it to at least 2--3$\times$ the interval.
    \item Network connectivity issues between probe and hub.
    \item Hub was restarted and the probe hasn't reported yet.
\end{itemize}

\subsection{BlackBox Reset}

\textbf{Symptom:} Probe logs ``Identity or hub key changed. Reinitializing.''

\textbf{Cause:} The probe or hub identity key has changed since the BlackBox
was last written to. Queued reports are cryptographically invalid.

\textbf{Resolution:} This is automatic. The BlackBox database is deleted and
recreated. Queued reports from the old key are lost (by design).

\subsection{TLS Certificate Mismatch}

\textbf{Symptom:} Probe fails to connect over HTTPS.

\textbf{Cause:} The hub's TLS certificate was regenerated but the probe was
not rebuilt with the new certificate.

\textbf{Resolution:}
\begin{enumerate}
    \item Copy the new \texttt{hub\_tls\_cert.pem} to the probe's build
          directory.
    \item Rebuild and redeploy the probe.
\end{enumerate}

\subsection{ICMP Ping Failures on Linux}

\textbf{Symptom:} All ping checks fail with ``Permission denied''.

\textbf{Cause:} Raw sockets require elevated privileges on Linux.

\textbf{Resolution:} Either run the probe as root or grant the capability:

\begin{lstlisting}
sudo setcap cap_net_raw+ep /path/to/probe
\end{lstlisting}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART II — TECHNICAL REFERENCE
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{Technical Reference}
\setcounter{section}{0}

% ============================================================
\section{Security Model}
% ============================================================

Blentinel implements a three-layer security model. Each layer operates
independently, so a failure in one layer does not compromise the others.

\subsection{Layer 1 --- Application-Layer Encryption}

Every report is encrypted at the application level before transmission:

\begin{enumerate}
    \item The probe generates an \textbf{ephemeral X25519 keypair} for each
          report (providing perfect forward secrecy).
    \item An ECDH shared secret is derived:
          $\texttt{shared} = \texttt{ephemeral\_secret} \otimes
          \texttt{hub\_public\_key}$.
    \item The report JSON is encrypted with
          \textbf{ChaCha20-Poly1305} (AEAD):
    \begin{itemize}
        \item Key: the 32-byte shared secret.
        \item Nonce: 12 random bytes (generated per report).
        \item Output: ciphertext + 16-byte authentication tag.
    \end{itemize}
    \item The final binary payload is:
\end{enumerate}

\begin{lstlisting}
[32 bytes: ephemeral X25519 public key]
[12 bytes: nonce]
[N  bytes: ciphertext + auth tag]
\end{lstlisting}

Total overhead: 44 bytes + 16-byte auth tag per report.

\subsection{Layer 2 --- Digital Signatures}

Every report is signed with the probe's \textbf{Ed25519} private key before
encryption:

\begin{enumerate}
    \item A \texttt{SignableReport} is constructed (the full report minus the
          \texttt{signature} and \texttt{ephemeral\_public\_key} fields).
    \item The signable report is serialised to canonical JSON bytes.
    \item A 64-byte Ed25519 signature is computed over those bytes.
    \item The hub verifies the signature using the probe's public key
          (registered in the whitelist).
\end{enumerate}

\textbf{Note:} Field order in \texttt{SignableReport} is part of the signed
canonical representation. Reordering fields is a breaking protocol change.

\subsection{Layer 3 --- Transport-Layer Security (Optional)}

When enabled, TLS provides an additional layer:

\begin{itemize}
    \item ECDSA P-256 self-signed certificate (auto-generated).
    \item Certificate pinning at compile time: probes embed the hub's
          certificate via \texttt{include\_bytes!}.
    \item No system CA dependency --- prevents MITM even with compromised
          certificate authorities.
\end{itemize}

\subsection{Zero-Trust Whitelisting}

The hub rejects any probe whose Ed25519 public key is not explicitly listed in
the \texttt{[[probes]]} section of \texttt{blentinel\_hub.toml}. The whitelist
is checked after decryption but before the report is persisted.

\subsection{Web Interface Authentication}

The hub's web interface is protected by token-based session authentication:

\begin{enumerate}
    \item On first run, a 32-character hexadecimal admin token is generated
          from the OS random source and saved to \texttt{hub\_auth.token}.
    \item Administrators log in by submitting this token via
          \texttt{POST /api/login}.
    \item A session cookie (\texttt{blentinel\_session}) is set with
          \texttt{HttpOnly}, \texttt{SameSite=Lax}, 24-hour expiry.
    \item All dashboard and admin API endpoints require a valid session.
    \item Probe endpoints (\texttt{/api/handshake}, \texttt{/api/report}) do
          not require a session --- they authenticate via cryptography.
\end{enumerate}

% ============================================================
\section{Report Format}
% ============================================================

\subsection{StatusReport Structure}

The complete JSON report sent by the probe (before encryption):

\begin{lstlisting}
{
  "probe_id": "2372960776...fbc4d5",
  "company_id": "IT_CLIENT_ALPHA",
  "hostname": "office-pc-01",
  "site": "main",
  "timestamp": "2025-02-14T15:30:45.123456Z",
  "interval_seconds": 60,
  "resources": [
    {
      "name": "Router",
      "resource_type": "ping",
      "target": "192.168.1.1",
      "status": "Up",
      "message": "Responded in 2ms",
      "latency_ms": 2,
      "metric_value": null,
      "metric_unit": null
    }
  ],
  "signature": [/* 64 bytes */],
  "ephemeral_public_key": [/* 32 bytes */]
}
\end{lstlisting}

\subsection{ResourceStatus Fields}

\begin{longtable}{@{} l l l p{5cm} @{}}
\toprule
\textbf{Field} & \textbf{Type} & \textbf{Optional} & \textbf{Description} \\
\midrule
\texttt{name} & String & No & Resource name from configuration. \\
\texttt{resource\_type} & Enum & No & One of: \texttt{ping}, \texttt{http},
    \texttt{tcp}, \texttt{local\_disk}, \texttt{local\_cpu},
    \texttt{local\_mem}, \texttt{local\_uptime}. \\
\texttt{target} & String & No & Target address or mount point. \\
\texttt{status} & Enum & No & \texttt{Up} or \texttt{Down}. \\
\texttt{message} & String & No & Descriptive status message. \\
\texttt{latency\_ms} & u64 & Yes & Latency in milliseconds (ping, tcp). \\
\texttt{metric\_value} & f64 & Yes & Metric value (CPU\%, memory\%, disk\%).
    \\
\texttt{metric\_unit} & String & Yes & Unit: \texttt{"\%"} or
    \texttt{"seconds"}. \\
\bottomrule
\end{longtable}

\subsection{Encrypted Payload Format}

The binary payload sent to \texttt{POST /api/report}:

\begin{lstlisting}
Bytes  0-31:   Ephemeral X25519 public key (32 bytes)
Bytes 32-43:   ChaCha20-Poly1305 nonce     (12 bytes)
Bytes 44+:     Ciphertext + auth tag        (N + 16 bytes)
\end{lstlisting}

Total size: 44 bytes of overhead + ciphertext length (approximately equal to
the JSON report size) + 16-byte authentication tag.

% ============================================================
\section{API Reference}
% ============================================================

All API endpoints are served by the hub's Axum router.

\subsection{Probe Endpoints (No Session Required)}

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Description} \\
\midrule
\texttt{GET} & \texttt{/api/handshake} & Returns the hub's X25519 public key
    (hex-encoded). Used by probes during initial key exchange. \\
\texttt{POST} & \texttt{/api/report} & Receives an encrypted probe report.
    Content-Type: \texttt{application/octet-stream}. Returns 200 on success,
    401 if decryption fails, 403 if the probe is not whitelisted or signature
    verification fails. \\
\bottomrule
\end{longtable}

\subsection{Authentication Endpoints}

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Description} \\
\midrule
\texttt{POST} & \texttt{/api/login} & Accepts the admin token in the request
    body. Sets a session cookie on success (24-hour expiry). \\
\texttt{POST} & \texttt{/api/logout} & Destroys the current session and
    clears the cookie. \\
\bottomrule
\end{longtable}

\subsection{Dashboard Endpoints (Session Required)}

\begin{longtable}{@{} l l p{5.5cm} @{}}
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Description} \\
\midrule
\texttt{GET} & \texttt{/api/dashboard/companies} & Returns all companies with
    probe counts, device status, and last report timestamp. \\
\texttt{GET} & \texttt{/api/company/\{id\}/probes} & Returns all probes for a
    company with status and device counts. \\
\texttt{GET} & \texttt{/api/company/\{id\}/uptime} & Returns uptime history
    buckets. Accepts \texttt{?range=} query parameter: \texttt{24h} (default),
    \texttt{7d}, \texttt{30d}, \texttt{all}. \\
\texttt{GET} & \texttt{/api/probe/\{id\}/devices} & Returns the latest
    resource statuses for a probe. \\
\bottomrule
\end{longtable}

\subsection{Admin Endpoints (Session Required)}

\begin{longtable}{@{} l l p{5cm} @{}}
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Description} \\
\midrule
\texttt{GET} & \texttt{/api/admin/companies} & Lists all known company IDs.
    \\
\texttt{GET} & \texttt{/api/admin/probes} & Lists all probes with status and
    timestamps. \\
\texttt{POST} & \texttt{/api/admin/delete-company-data} & Deletes all reports
    and resource statuses for a company. JSON body:
    \texttt{\{"company\_id": "..."\}}. \\
\texttt{POST} & \texttt{/api/admin/delete-probe-data} & Deletes all reports
    for a probe (keeps heartbeat). JSON body:
    \texttt{\{"probe\_id": "..."\}}. \\
\texttt{POST} & \texttt{/api/admin/remove-probe} & Removes a probe entirely:
    deletes data, heartbeat, and all traces. \\
\texttt{GET} & \texttt{/api/admin/storage-info} & Returns current database
    size, warning threshold, and last archive info. \\
\texttt{GET} & \texttt{/api/admin/archives} & Lists all archive records. \\
\texttt{POST} & \texttt{/api/admin/archive} & Creates a new archive. Optional
    JSON body: \texttt{\{"cutoff\_days": 90\}}. \\
\bottomrule
\end{longtable}

\subsection{Alert Silence Endpoints (Session Required)}

\begin{longtable}{@{} l l p{5cm} @{}}
\toprule
\textbf{Method} & \textbf{Path} & \textbf{Description} \\
\midrule
\texttt{GET} & \texttt{/api/silences} & Lists all active alert silences. \\
\texttt{POST} & \texttt{/api/silence} & Creates a silence. JSON body:
    \texttt{\{"resource\_key": "...", "reason": "...",
    "duration\_hours": 24\}}. Omit \texttt{duration\_hours} for indefinite.
    \\
\texttt{POST} & \texttt{/api/silence/delete} & Deletes a silence. JSON body:
    \texttt{\{"silence\_id": 1\}}. \\
\bottomrule
\end{longtable}

\subsection{Archive Viewer Endpoints (Session Required)}

Read-only endpoints for browsing historical archive databases:

\begin{longtable}{@{} l p{9cm} @{}}
\toprule
\textbf{Method \& Path} & \textbf{Description} \\
\midrule
\texttt{GET /api/archive/\{id\}/companies} & Dashboard view of archived data.
    \\
\texttt{GET /api/archive/\{id\}/company/\{cid\}/probes} & Probes in an
    archived company. \\
\texttt{GET /api/archive/\{id\}/company/\{cid\}/uptime} & Uptime history from
    archive. \\
\texttt{GET /api/archive/\{id\}/probe/\{pid\}/devices} & Device statuses from
    archive. \\
\bottomrule
\end{longtable}

% ============================================================
\section{Uptime History Bucketing}
% ============================================================

The uptime endpoint uses adaptive time bucketing to keep the data point count
bounded:

\begin{longtable}{@{} l l l l @{}}
\toprule
\textbf{Range} & \textbf{Window} & \textbf{Bucket Size} &
    \textbf{Approx.\ Points} \\
\midrule
\texttt{24h} & Last 24 hours & 15 minutes & $\sim$96 \\
\texttt{7d}  & Last 7 days   & 1 hour     & $\sim$168 \\
\texttt{30d} & Last 30 days  & 4 hours    & $\sim$180 \\
\texttt{all} & All data      & 1 day      & $\leq$ retention days \\
\bottomrule
\end{longtable}

% ============================================================
\section{Database Schema}
% ============================================================

\subsection{Hub Database (\texttt{blentinel.db})}

The hub uses SQLite in WAL (Write-Ahead Logging) mode with a 5-second busy
timeout, ensuring readers never block writers.

\subsubsection{Core Tables}

\paragraph{\texttt{reports}} Stores one row per received probe report.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{id} & INTEGER PK & Auto-incrementing primary key. \\
\texttt{probe\_id} & TEXT & Probe's Ed25519 public key (hex). \\
\texttt{company\_id} & TEXT & Company identifier. \\
\texttt{hostname} & TEXT & Probe's hostname. \\
\texttt{site} & TEXT & Logical site name. \\
\texttt{timestamp} & DATETIME & Report timestamp (UTC). \\
\texttt{interval\_seconds} & INTEGER & Probe's reporting interval. \\
\texttt{received\_at} & DATETIME & When the hub received the report. \\
\bottomrule
\end{longtable}

\paragraph{\texttt{resource\_statuses}} Stores one row per resource per
report.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{id} & INTEGER PK & Auto-incrementing primary key. \\
\texttt{report\_id} & INTEGER FK & References \texttt{reports(id)}. \\
\texttt{name} & TEXT & Resource name. \\
\texttt{resource\_type} & TEXT & Resource type string. \\
\texttt{target} & TEXT & Target address. \\
\texttt{status} & TEXT & ``Up'' or ``Down''. \\
\texttt{message} & TEXT & Descriptive status message. \\
\texttt{latency\_ms} & INTEGER & Latency in milliseconds (nullable). \\
\texttt{metric\_value} & REAL & Metric value (nullable). \\
\texttt{metric\_unit} & TEXT & Metric unit (nullable). \\
\bottomrule
\end{longtable}

\subsubsection{Probe Liveness}

\paragraph{\texttt{probe\_heartbeats}} Tracks probe liveness. Updated on
every successful report; a background task (running every 10 seconds) flips
the status to ``expired'' when \texttt{last\_seen\_at} exceeds the configured
timeout.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{probe\_id} & TEXT PK & Probe's Ed25519 public key. \\
\texttt{company\_id} & TEXT & Company identifier. \\
\texttt{status} & TEXT & ``active'' or ``expired''. \\
\texttt{last\_seen\_at} & DATETIME & Timestamp of last report. \\
\texttt{first\_seen\_at} & DATETIME & Timestamp of first report. \\
\bottomrule
\end{longtable}

Probes automatically resurrect (flip back to ``active'') when their next
report arrives.

\subsubsection{Archival}

\paragraph{\texttt{archives}} Metadata for each archive database created.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{id} & INTEGER PK & Auto-incrementing primary key. \\
\texttt{filename} & TEXT UNIQUE & Archive filename. \\
\texttt{created\_at} & DATETIME & When the archive was created. \\
\texttt{cutoff\_date} & DATETIME & Data older than this was archived. \\
\texttt{size\_mb} & INTEGER & Archive file size in megabytes. \\
\bottomrule
\end{longtable}

\subsubsection{Alert State Tracking}

\paragraph{\texttt{alert\_states}} Tracks the last known status of each
resource for alert deduplication. Alerts are only sent on state transitions
(e.g.\ Up $\to$ Down), not on every report.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{resource\_key} & TEXT PK & Unique resource identifier. \\
\texttt{last\_status} & TEXT & Last known status. \\
\texttt{last\_alert\_sent\_at} & DATETIME & When the last alert was sent
    (nullable). \\
\bottomrule
\end{longtable}

\paragraph{\texttt{alert\_silences}} Manages alert suppression at various
scopes.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{id} & INTEGER PK & Auto-incrementing primary key. \\
\texttt{scope\_type} & TEXT & Scope: ``company'', ``probe'', or ``resource''.
    \\
\texttt{scope\_id} & TEXT & Identifier within the scope. \\
\texttt{reason} & TEXT & Human-readable reason for the silence. \\
\texttt{created\_at} & DATETIME & When the silence was created. \\
\texttt{expires\_at} & DATETIME & Optional expiry (NULL = indefinite). \\
\bottomrule
\end{longtable}

Unique constraint on \texttt{(scope\_type, scope\_id)}.

\subsubsection{Indexes}

\begin{lstlisting}
idx_reports_probe_id        ON reports(probe_id)
idx_reports_company_id      ON reports(company_id)
idx_reports_timestamp       ON reports(timestamp)
idx_resource_statuses_report_id ON resource_statuses(report_id)
idx_heartbeats_company      ON probe_heartbeats(company_id)
idx_archives_created        ON archives(created_at)
idx_silences_expires        ON alert_silences(expires_at)
\end{lstlisting}

\subsection{Probe BlackBox Database (\texttt{blentinel\_blackbox.db})}

The BlackBox is a local store-and-forward queue used when the hub is
unreachable.

\paragraph{\texttt{pending\_reports}} Queued encrypted payloads.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{id} & INTEGER PK & Auto-incrementing primary key. \\
\texttt{payload} & BLOB & Encrypted binary payload (same format as
    \texttt{POST /api/report}). \\
\texttt{created\_at} & DATETIME & When the report was queued. \\
\bottomrule
\end{longtable}

\paragraph{\texttt{meta}} Identity fingerprints for integrity validation.

\begin{longtable}{@{} l l p{6cm} @{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Description} \\
\midrule
\texttt{key} & TEXT PK & Metadata key. \\
\texttt{value} & TEXT & Metadata value. \\
\bottomrule
\end{longtable}

Stored keys:
\begin{itemize}
    \item \texttt{probe\_public\_key} --- current probe's Ed25519 public key.
    \item \texttt{hub\_public\_key} --- current hub's X25519 public key.
\end{itemize}

On startup, the BlackBox compares these fingerprints with the current keys. If
either has changed, the database is deleted because queued payloads would be
cryptographically invalid (encrypted with the old hub key or signed with the
old probe key).

% ============================================================
\section{Alert Deduplication Internals}
% ============================================================

Alerts are deduplicated using the \texttt{alert\_states} table. An alert is
only sent on state transitions, not on every report. This prevents alert
storms when a resource remains down across multiple reporting cycles.

% ============================================================
\section{Probe Expiry Mechanism}
% ============================================================

A background task on the hub runs every 10 seconds, scanning for probes whose
\texttt{last\_seen\_at} exceeds the configured \texttt{probe\_timeout\_secs}.
These probes are marked ``expired''. When an expired probe sends its next
report, it is automatically resurrected to ``active'' status.

% ============================================================
\section{Background Tasks}
% ============================================================

The hub runs several concurrent background tasks:

\begin{itemize}
    \item \textbf{Probe expiry checker} --- every 10 seconds.
    \item \textbf{Database size monitor} --- every 10 minutes (if retention is
          enabled).
    \item \textbf{Configuration file watcher} --- continuous filesystem
          monitoring.
    \item \textbf{Alert evaluation} --- spawned asynchronously after each
          report and each expiry check.
\end{itemize}

% ============================================================
\section{Error Handling}
% ============================================================

\subsection{Configuration Errors}

Both hub and probe define a \texttt{ConfigError} enum with four variants:

\begin{longtable}{@{} l p{8cm} @{}}
\toprule
\textbf{Variant} & \textbf{Meaning} \\
\midrule
\texttt{NotFound} & Configuration file does not exist. \\
\texttt{Io} & I/O error reading the file. \\
\texttt{Invalid} & TOML parse error. \\
\texttt{Validation} & Semantic validation failed (e.g.\ interval = 0,
    empty URL, invalid hex key). \\
\bottomrule
\end{longtable}

During hot reload, configuration errors are logged and the previous valid
configuration is retained.

\subsection{Hub Error Responses}

\begin{longtable}{@{} l l p{5cm} @{}}
\toprule
\textbf{HTTP Status} & \textbf{Condition} & \textbf{Meaning} \\
\midrule
200 & --- & Report accepted and saved. \\
400 & Bad Request & Decrypted payload is not valid JSON. \\
401 & Unauthorised & Decryption failed (wrong key) or missing/invalid
    session cookie. \\
403 & Forbidden & Probe not in whitelist or signature verification failed. \\
500 & Internal Error & Database or server error. \\
\bottomrule
\end{longtable}

\subsection{Logging}

\begin{itemize}
    \item \texttt{--verbose}: Operational events (reports received, probes
          expired, database operations, handshake progress).
    \item \texttt{--debug}: Full diagnostic output including cleartext
          payloads (probe) and detailed request/response logging (hub).
    \item Errors are always printed to \texttt{stderr}.
\end{itemize}

\textcolor{warnorange}{\textbf{Warning:}} Never use \texttt{--debug} in
production. It logs cleartext report data, which defeats the purpose of
encryption.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PART III — DEVELOPER DOCUMENTATION
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\part{Developer Documentation}
\setcounter{section}{0}

% ============================================================
\section{Workspace Layout}
% ============================================================

The project uses a Cargo workspace with shared release profile optimisations:

\begin{lstlisting}
[workspace]
members = ["probe", "hub", "common", "blentinelmake"]

[profile.release]
opt-level = "z"      # Optimise for binary size
lto = true           # Link-time optimisation
panic = "abort"      # No unwinding overhead

[profile.wasm-release]
inherits = "release"
codegen-units = 1    # Single codegen unit for WASM
\end{lstlisting}

% ============================================================
\section{Building from Source}
% ============================================================

\subsection{Prerequisites}

\begin{itemize}
    \item Rust toolchain (stable, edition 2024 for probe/common, 2021 for
          hub/blentinelmake).
    \item \texttt{cargo-leptos} for building the hub:
          \texttt{cargo install cargo-leptos}.
    \item WASM target: \texttt{rustup target add wasm32-unknown-unknown}.
\end{itemize}

\subsection{Using blentinelmake}

Build the build tool first:

\begin{lstlisting}
cargo build -p blentinelmake --release
\end{lstlisting}

\paragraph{Interactive Mode} Run without arguments for a menu-driven
interface:

\begin{lstlisting}
./target/release/blentinelmake
\end{lstlisting}

The interactive mode presents choices for component (hub/probe), action
(build/publish/clean), target triple, and release/debug mode.

\paragraph{CLI Mode}

\begin{lstlisting}
blentinelmake <component> <action> [OPTIONS]
\end{lstlisting}

Examples:

\begin{lstlisting}
blentinelmake hub build --release
blentinelmake hub publish
blentinelmake probe build --release
blentinelmake probe publish --target x86_64-unknown-linux-musl
blentinelmake probe clean
\end{lstlisting}

\subsection{Manual Build}

\paragraph{Hub}

\begin{lstlisting}
cargo leptos build --release
\end{lstlisting}

This produces both the server binary (SSR) and the WASM client bundle.

\paragraph{Probe}

\begin{lstlisting}
cargo build -p probe --release
\end{lstlisting}

% ============================================================
\section{Publishing}
% ============================================================

The \texttt{publish} action builds in release mode and packages artefacts
into a distributable directory:

\begin{itemize}
    \item \texttt{publish/hub/app/} --- hub binary, sample config, service
          scripts, SHA256SUM.
    \item \texttt{publish/probe/<target>/app/} --- probe binary, sample
          config, service scripts, SHA256SUM, \texttt{hub\_cert.pem} (if
          available).
\end{itemize}

ZIP archives are also created for convenience.

\subsection{SHA-256 Integrity Verification}

Each published package includes a \texttt{SHA256SUM} file:

\begin{lstlisting}
<64-char-hex-hash>  <filename>
\end{lstlisting}

Verify on Linux/macOS:

\begin{lstlisting}
sha256sum -c SHA256SUM
\end{lstlisting}

Verify on Windows (PowerShell):

\begin{lstlisting}
$hash = (Get-FileHash hub.exe -Algorithm SHA256).Hash
$expected = (Get-Content SHA256SUM).Split()[0]
if ($hash -eq $expected) { "OK" } else { "FAILED" }
\end{lstlisting}

% ============================================================
\section{Cross-Compilation}
% ============================================================

The \textbf{Hub cannot be cross-compiled} due to Leptos WASM constraints. It
must be built natively on the target platform.

The \textbf{Probe can be cross-compiled} to any supported target.

\subsection{Supported Targets}

\begin{longtable}{@{} l l @{}}
\toprule
\textbf{Target Triple} & \textbf{Platform} \\
\midrule
\texttt{x86\_64-pc-windows-msvc} & 64-bit Windows \\
\texttt{x86\_64-unknown-linux-gnu} & 64-bit Linux (glibc) \\
\texttt{x86\_64-unknown-linux-musl} & 64-bit Linux (static, musl) \\
\texttt{aarch64-unknown-linux-gnu} & 64-bit ARM Linux (glibc) \\
\texttt{aarch64-unknown-linux-musl} & 64-bit ARM Linux (static, musl) \\
\bottomrule
\end{longtable}

\subsection{Setup}

Install the required Rust targets:

\begin{lstlisting}
rustup target add x86_64-unknown-linux-gnu
rustup target add x86_64-unknown-linux-musl
rustup target add aarch64-unknown-linux-gnu
rustup target add aarch64-unknown-linux-musl
\end{lstlisting}

Optionally install \texttt{cargo-zigbuild} for easier cross-compilation
without Docker:

\begin{lstlisting}
cargo install cargo-zigbuild
\end{lstlisting}

\subsection{Static Linking (musl Targets)}

The \texttt{musl} targets produce fully static binaries with no runtime
dependencies on glibc or other system libraries. These binaries work on any
Linux system with the matching architecture, making them ideal for
distribution. The trade-off is a larger binary size.

\subsection{Binary Stripping}

During the \texttt{publish} action, debug symbols are automatically stripped
from non-Windows binaries, reducing binary size by approximately 60\%. Windows
executables are not stripped (a separate \texttt{.pdb} file may be available).

% ============================================================
\section{Web Interface Architecture}
% ============================================================

The hub serves a full-stack web application built with Leptos 0.8.

\subsection{Architecture}

\begin{itemize}
    \item \textbf{Server-Side Rendering (SSR):} Pages are rendered on the
          server using Axum, ensuring fast initial page loads and
          functionality without JavaScript.
    \item \textbf{WebAssembly Hydration:} After initial render, the browser
          loads a WASM bundle that ``hydrates'' the page, enabling interactive
          behaviour.
    \item \textbf{Styling:} SCSS compiled to CSS.
\end{itemize}

\subsection{Pages}

\begin{itemize}
    \item \textbf{Login:} Token input for admin authentication.
    \item \textbf{Dashboard:} Card grid showing all companies with aggregate
          status (total/active/expired probes, resources up/down, last
          report).
    \item \textbf{Company Detail:} List of all probes for a company with
          status indicators, hostname, site, and last seen timestamp.
    \item \textbf{Probe Detail:} All monitored resources for a probe with
          status, latency, metrics, and uptime graphs.
    \item \textbf{Admin Panel:} Database storage information, archive
          management, probe and company data deletion.
    \item \textbf{Archive Viewer:} Browse historical archive databases using
          the same dashboard/company/probe views.
\end{itemize}

\subsection{Deployed Mode Detection}

The hub auto-detects deployed mode: if a \texttt{pkg/} directory exists in
the working directory, it serves assets from \texttt{./} instead of the
compile-time path (\texttt{target/site/}). This allows the published binary
to function without the full build tree.

% ============================================================
\section{Source Directory Structure}
% ============================================================

\begin{lstlisting}
blentinel/
  Cargo.toml                   # Workspace manifest
  common/
    Cargo.toml
    src/
      lib.rs                   # Crate root
      models.rs                # Shared data models
  probe/
    Cargo.toml
    src/
      main.rs                  # Entry point
      args.rs                  # CLI argument parsing
      config.rs                # Configuration loading
      identity.rs              # Ed25519 key management
      crypto.rs                # Signing + encryption
      transport.rs             # HTTP(S) report delivery
      monitor.rs               # Resource polling loop
      checks.rs                # Check implementations
      storage.rs               # BlackBox database
      hot_reload.rs            # Config file watcher
      tls.rs                   # Certificate pinning
  hub/
    Cargo.toml
    src/
      main.rs                  # Entry point + Axum router
      lib.rs                   # Leptos app registration
      app.rs                   # Leptos frontend components
      args.rs                  # CLI argument parsing
      config.rs                # Configuration loading
      identity.rs              # X25519 key management
      crypto.rs                # Decryption + verification
      auth.rs                  # Session authentication
      api.rs                   # REST API handlers
      db.rs                    # Database queries
      db/types.rs              # SQLx type adapters
      hot_reload.rs            # Config file watcher
      tls.rs                   # Certificate generation
      archive.rs               # Archival system
      archive/engine.rs        # Archive creation logic
      archive/monitor.rs       # DB size monitor
      archive_viewer.rs        # Archive read-only API
      alerts.rs                # Alert system entry
      alerts/engine.rs         # Alert evaluation logic
      alerts/state.rs          # Alert state management
      alerts/silence.rs        # Silence management
      alerts/email.rs          # SMTP e-mail sending
    style/
      main.scss                # Frontend styles
    public/                    # Static assets
  blentinelmake/
    Cargo.toml
    src/
      main.rs                  # Build tool
  manual/
    manual.tex                 # This document
\end{lstlisting}

% ============================================================
\section{Glossary}
% ============================================================

\begin{longtable}{@{} l p{9cm} @{}}
\toprule
\textbf{Term} & \textbf{Definition} \\
\midrule
AEAD & Authenticated Encryption with Associated Data. ChaCha20-Poly1305 is
    the AEAD cipher used by Blentinel. \\
BlackBox & The probe's local SQLite database that queues encrypted reports
    when the hub is unreachable. \\
ECDH & Elliptic-Curve Diffie--Hellman key exchange. Used with X25519 to
    derive a shared secret from ephemeral and static keys. \\
Ed25519 & Edwards-curve Digital Signature Algorithm. Used for probe identity
    and report signing. \\
Hub & The central server that receives, authenticates, decrypts, and stores
    monitoring data. \\
Hydration & The process by which a server-rendered HTML page is ``activated''
    by client-side JavaScript/WASM for interactivity. \\
MSP & Managed Service Provider. An IT services company that manages client
    infrastructure remotely. \\
PFS & Perfect Forward Secrecy. Ensured by generating ephemeral X25519 keys
    per report. \\
Probe & A lightweight agent deployed in a client network that collects and
    reports monitoring data. \\
SSR & Server-Side Rendering. The hub renders full HTML on the server before
    sending it to the browser. \\
WAL & Write-Ahead Logging. SQLite journal mode that allows concurrent reads
    and writes. \\
X25519 & Elliptic-curve key agreement protocol based on Curve25519. Used for
    asymmetric encryption key exchange. \\
\bottomrule
\end{longtable}

\end{document}
